Estimating demographic bias on tests of children's early vocabulary

### 2022-01-17 paper/cogsci.Rmd
wrapping up results

# regression - report coefficients again after we cut some items?
require(lmerTest)
d_demo$eth_group = factor(d_demo$eth_group, levels=c("Nonwhite","White"))
d_demo$sex = factor(d_demo$sex, levels=c("Male","Female"))
d_demo$ses_group = factor(d_demo$ses_group, levels=c("low","high"))
d_demo <- d_demo %>% mutate(age_sc = scale(age, scale=F))
eth_reg <- summary(lm(ability ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 white-, age*white+
ses_reg <- summary(lm(ability ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg <- summary(lm(ability ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

@@

coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
fscores_1pl <- tibble(data_id = d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(data_id), 
                             ability = fscores(m0, method = "MAP")[,1])

save(file = here("data/eng_Ws_prod_mod_1pl_pruned.Rds"), "m0", "coefs_1pl", "fscores_1pl")
```

@@

# join with d_demo, do the demo regressions again..
```

We identified 76 CDI items that function quite differently across demographic groups, and chose to prune the 15 items with large difficulty differences for both SES and race.

```{r demo-ability, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.3, fig.height=5.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Density plots of language ability for children aged 23-25 months, based on the group models."}
# show density plots of child ability based on group models

@@

# sex
sex <- d_demo %>% filter(!is.na(sex), production!=0) %>%
  mutate(theta_sex = fscores(mod_intuitive_sex)) 
# cor.test(eth$production, eth$theta_eth) # .98

@@

# SES
ses <- d_demo %>% filter(!is.na(ses_group), production!=0) %>%
  mutate(theta_ses = fscores(mod_intuitive_ses)) %>%
  filter(age<26, age>22) %>%
  rename(SES=ses_group)
# cor.test(eth$production, eth$theta_eth) # .98

@@

# ethnicity
eth <- d_demo %>% filter(!is.na(eth_group), production!=0) %>%
  mutate(theta_eth = fscores(mod_intuitive_eth)) %>%
  filter(age<26, age>22) %>%
  mutate(Race=eth_group)
# cor.test(eth$production, eth$theta_eth) # .98

>>>>

# regression - report coefficients again after we cut some items?
d_demo$eth_group = factor(d_demo$eth_group, levels=c("Nonwhite","White"))
d_demo$sex = factor(d_demo$sex, levels=c("Male","Female"))
d_demo$ses_group = factor(d_demo$ses_group, levels=c("low","high"))
d_demo <- d_demo %>% mutate(age_sc = scale(age, scale=F))
eth_reg <- summary(lm(ability ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 white-, age*white+
ses_reg <- summary(lm(ability ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg <- summary(lm(ability ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

@@

coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
fscores_1pl <- tibble(data_id = d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(data_id), 
                             ability_pr = fscores(m0, method = "MAP")[,1])
fscores_1pl$data_id = fscores_1pl$data_id$data_id
save(file = here("data/eng_Ws_prod_mod_1pl_pruned.Rds"), "m0", "coefs_1pl", "fscores_1pl")
```

@@

d_demo <- d_demo %>% left_join(fscores_1pl)
# join with d_demo, do the demo regressions again..

eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 
ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

#white=.48
#highSES=0.20
#female=0.57
```

We identified 76 CDI items that function quite differently across demographic groups, and chose to prune the 15 items with large difficulty differences for both SES and race. 
After pruning these 2.2% of CDI:WS items, regressions with each demographic variable (and age, centered) showed smaller effects of SES ($\beta=0.20$ vs. $\beta=0.23$) and race ($\beta=0.48$ vs. $\beta=0.50$), while the effect of sex remained the same.

```{r demo-ability, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.3, fig.height=5.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Density plots of language ability for children aged 23-25 months using the pruned model."}

@@

# sex
sex <- d_demo %>% filter(!is.na(sex), production!=0) %>%
  mutate(theta_sex = ability_pr)
  #mutate(theta_sex = fscores(mod_intuitive_sex)) 
# cor.test(eth$production, eth$theta_eth) # .98

@@

# SES
ses <- d_demo %>% filter(!is.na(ses_group), production!=0) %>%
  #mutate(theta_ses = fscores(mod_intuitive_ses)) %>%
  mutate(theta_ses = ability_pr) %>%
  filter(age<26, age>22) %>%
  rename(SES=ses_group)
# cor.test(eth$production, eth$theta_eth) # .98

@@

# ethnicity
eth <- d_demo %>% filter(!is.na(eth_group), production!=0) %>%
  #mutate(theta_eth = fscores(mod_intuitive_eth)) %>%
  mutate(theta_eth = ability_pr) %>%
  filter(age<26, age>22) %>%
  mutate(Race=eth_group)
# cor.test(eth$production, eth$theta_eth) # .98
